# Suggested Answers {#SuggestedAnswers}

This Appendix contains suggested answers for the exercises in the book.

## SQL and R {#SqlAndRanswers}

**Exercise \@ref(exr:exercisePeopleCount)**

To compute the number of people we can simply query the PERSON table:

```{r eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(*) AS person_count
FROM @cdm.person;"

renderTranslateQuerySql(connection, sql, cdm = "main")
```
```{r echo=FALSE,message=FALSE,eval=TRUE}
cat("  PERSON_COUNT
1         2694
")
```

**Exercise \@ref(exr:exerciseCelecoxibUsers)**

To compute the number of people with at least one prescription of celecoxib, we can query the DRUG_EXPOSURE table. To find all drugs containing the ingredient celecoxib, we join to the CONCEPT_ANCESTOR and CONCEPT tables:

```{r eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(DISTINCT(person_id)) AS person_count
FROM @cdm.drug_exposure
INNER JOIN @cdm.concept_ancestor
  ON drug_concept_id = descendant_concept_id
INNER JOIN @cdm.concept ingredient
  ON ancestor_concept_id = ingredient.concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S';"

renderTranslateQuerySql(connection, sql, cdm = "main")
```
```{r echo=FALSE,message=FALSE,eval=TRUE}
cat("  PERSON_COUNT
1         1844
")
```

Note that we use `COUNT(DISTINCT(person_id))` to find the number of distinct persons, considering that a person might have more than one prescription. Also note that we use the `LOWER` function to make our search for "celecoxib" case-insensitive.

Alternatively, we can use the DRUG_ERA table, which is already rolled up to the ingredient level:

```{r eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)

sql <- "SELECT COUNT(DISTINCT(person_id)) AS person_count
FROM @cdm.drug_era
INNER JOIN @cdm.concept ingredient
  ON drug_concept_id = ingredient.concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S';"

renderTranslateQuerySql(connection, sql, cdm = "main")
```
```{r echo=FALSE,message=FALSE,eval=TRUE}
cat("  PERSON_COUNT
1         1844
")
```

**Exercise \@ref(exr:exerciseGiBleedsDuringCelecoxib)**

To compute the number of diagnoses during exposure we extend our previous query by joining to the CONDITION_OCCURRENCE table. We join to the CONCEPT_ANCESTOR table to find all condition concepts that imply a gastrointestinal haemorrhage:

```{r eval=FALSE}
library(DatabaseConnector)
connection <- connect(connectionDetails)
sql <- "SELECT COUNT(*) AS diagnose_count
FROM @cdm.drug_era
INNER JOIN @cdm.concept ingredient
  ON drug_concept_id = ingredient.concept_id
INNER JOIN @cdm.condition_occurrence
  ON condition_start_date >= drug_era_start_date
    AND condition_start_date <= drug_era_end_date
INNER JOIN @cdm.concept_ancestor 
  ON condition_concept_id =descendant_concept_id
WHERE LOWER(ingredient.concept_name) = 'celecoxib'
  AND ingredient.concept_class_id = 'Ingredient'
  AND ingredient.standard_concept = 'S'
  AND ancestor_concept_id = 192671;"

renderTranslateQuerySql(connection, sql, cdm = "main")
```
```{r echo=FALSE,message=FALSE,eval=TRUE}
cat("  DIAGNOSE_COUNT
1         41
")
```

Note that in this case it is essential to use the DRUG_ERA table instead of the DRUG_EXPOSURE table, because drug exposures with the same ingredient can overlap, but drug eras can. This could lead to double counting. For example, imagine a person received two drug drugs containing celecoxib at the same time. This would be recorded as two drug exposures, so any diagnoses occurring during the exposure would be counted twice. The two exposures will be merged into a single non-overlapping drug era.


## Data Quality {#DataQualityanswers}

**Exercise \@ref(exr:exerciseRunAchilles)**

To run ACHILLES:

```{r eval=FALSE}
library(ACHILLES)
result <- achilles(connectionDetails,
                   cdmDatabaseSchema = "main",
                   resultsDatabaseSchema = "main",
                   sourceName = "Eunomia",
                   cdmVersion = "5.3.0")

```

**Exercise \@ref(exr:exerciseViewHeel)**

To run extract the ACHILLES Heel issue list:

```{r eval=FALSE}
heel <- fetchAchillesHeelResults(connectionDetails,
                                 resultsDatabaseSchema = "main")
View(heel)
```
